Design and Implementation of Probabilistic Programming Language Anglican

### 2016-11-27 paper/paper.tex

Similar to distributions, \textit{random processes}, related to
so called `exchangeable random procedures'\footnote{However, random sequences
generated by Anglican random processes are not required to be
exchangeable.}, are defined using
macro \texttt{defproc} and implement the
\texttt{anglican.runtime/{\linebreak[0]}random-process} protocol (that is,
an interface for these random-sequence objects). The protocol
has two methods --- \texttt{produce}, which returns a
distribution object, and \texttt{absorb}, which returns a new
random process object obtained by absorbing a value drawn from
the process. Here is a definition of the Poisson process:
\begin{lstlisting}[style=default]
(defproc PP
  "Poisson point process"
  [rate] [dist (exponential rate)]
  (produce [this] dist)
  (absorb [this value] this))
\end{lstlisting}
While distributions are random and non-functional,
random processes are deterministic and functional. Hence,
\texttt{produce} and \texttt{absorb} are called directly and do
not have corresponding special forms in Anglican:
\begin{itemize}
    \item Values produced by the process are sampled from the distribution
        returned by \texttt{produce}.
    \item When a value is observed from the distribution, a new
        process, with the observed value absorbed, is returned
        by \texttt{absorb}.
\end{itemize}
A common pattern of programming with processes in Anglican is
recursively computing a new process based on the observed value:
\begin{lstlisting}[style=default]
(loop [process (HawkesPointProcess)
       [value0 & values0] values] 
  (let [dist (produce process)]
    (observe dist value0)
    (recur (absorb process value0) values0)))
\end{lstlisting}
Unlike conventional implementations of random processes,
Anglican's random processes do not have mutable state.

>>>>

Similar to distributions, \textit{random processes}, related to
so called `exchangeable random procedures'\footnote{However, random sequences
generated by Anglican random processes are not required to be
exchangeable.}, are defined using
macro \texttt{defproc} and implement the
\texttt{anglican.runtime/{\linebreak[0]}random-process} protocol (that is,
an interface for these random-sequence objects). The protocol
has two methods --- \texttt{produce}, which returns a
distribution object, and \texttt{absorb}, which returns a new
random process object obtained by absorbing a value drawn from
the process. Here is a definition of a beta Bernouilli process:
\begin{lstlisting}[style=default]
(defproc beta-bernoulli
  "beta-Bernoulli process"
  [a b]
  (produce [this] (bernoulli (/ a (+ a b))))
  (absorb [this x]
    (case x
      0 (beta-bernoulli a (inc b))
      1 (beta-bernoulli (inc a) b))))
\end{lstlisting}
While distributions are random and non-functional,
random processes are deterministic and functional. Hence,
\texttt{produce} and \texttt{absorb} are called directly and do
not have corresponding special forms in Anglican:
\begin{itemize}
    \item Values produced by the process are sampled from the distribution
        returned by \texttt{produce}.
    \item When a value is observed from the distribution, a new
        process, with the observed value absorbed, is returned
        by \texttt{absorb}.
\end{itemize}
A common pattern of programming with processes in Anglican is
recursively computing a new process based on the observed value:
\begin{lstlisting}[style=default]
(loop [process (beta-bernoulli 1.0 1.0)
       [value0 & values0] values] 
  (let [dist (produce process)]
    (observe dist value0)
    (recur (absorb process value0) values0)))
\end{lstlisting}
Unlike conventional implementations of random processes,
Anglican's random processes do not have mutable state.
